---
title: "Course_Project_IS_part1"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(ggplot2)
```

The theoretical mean and standard deviation, respectively, of the exponential distribution are 1/lambda. These are to be compared to simulated values for a distribution of 1000 averages of 40 random exponentials:

```{r}
set.seed(1)

averages = 1000
n = 40
lambda = 0.2

mns = NULL
for (i in 1 : 1000) mns = c(mns, mean(rexp(n, lambda)))

sim_mean = mean(mns)
expected_mean = 1/lambda

var_mns = var(mns)
```

1. The sample mean is `r round(sim_mean,3)` and the theoretical mean is `r round(expected_mean)`.

2. The sample variance is `r round(var_mns,3)` and the theoretical variance is `r 1/40*(1/lambda)^2`.

3. The means are normally distributed, see figure below.

```{r, fig.width = 5, fig.height = 4, fig.align = "center"}


g = ggplot(data.frame(y = mns), aes(x = y)) +
    geom_histogram(aes(y = ..density..), binwidth = 0.2,
                   color = "black", fill = "white") +
    stat_function(fun = dnorm, geom = "line",
            args = list(mean = 1/lambda,
            sd = 1/(lambda*sqrt(n))),
            size = 1, col = "blue") + 
    ylab("Probability density") + 
    xlab("Mean") + 
    ggtitle("Comparison of simulated and theoretical\n exponential distribution")
g    
```